% Place acknowledgments below.
The completion of graduate school and this dissertation could not have been possible without the support from many people that I would like to thank now.

First off I would like to thank my PhD advisor Dr. Zi-Kui Liu. I am so grateful to him for his guidance and support throughout my graduate school career. He opened the doors for many wonderful opportunities and I learned more than I thought was imaginable. On top of that he continued to push me and help me try and learn new things. His constant positive attitude and outlook on life has been inspirational to be me and I will continue to benefit from this for the rest of my life.

I want to thank the Penn State Materials Science and Engineering Department. There have been multiple staff, faculty, and external advisory board members who have helped me, mentored me and opened the door to new opportunities. I would like to specifically thank my committee members Dr. Allison Beese, Dr. Kristen Fichthorn, Dr. Michael Manley and Dr. Susan Sinnott for their time, constructive criticism and advice throughout this journey.

I want to thank Dr. Ji-Cheng Zhao, Dr. Michael Manley, and Dr. Yong Du. Each of you gave me the opportunity to learn new things and expand my PhD experience through collaborations, and allowing me to visit your labs. 

I want to express my deepest thanks to my family. I am lucky to have multiple family members who support and love me including my parents, siblings, grandparents, aunts, uncles and cousins. There were many ups and downs in my educational pursuit and you were each there for me in multiple different ways and I am very grateful.

I would also like to thank the members of the Phases Research Lab past and present. It was an honor to get to work with so many wonderful people and to connect with previous members. I learned a lot from each of you and formed many lasting friendships amongst the group. Listing all of the members that positively impacted my life would take up too much space but I wanted to thank the members who served as mentors to me and helped me get my footing, Dr. Shun-Li Shang, Dr. Richard Otis, Dr. Greta Lindwall, and Dr. Xuan Liu. Additionally, I would like to thank the members who collaborated and helped me with my research, Dr. Yong-Jie Hu, Dr. Bi-Cheng Zhou, Austin Ross, Hongyeun Kim, Pinwen Guan, and Brandon Bocklund. 

Lastly, I want to thank all of my friends at Penn State, my significant other Tobias Frueh, and all of my CALPHAD friends. While there are many people here to name, I would like to especially thank Carl Arvhult, Chris Rotella, Dan Schulman, Florian Tang, Lyndsey Denis, Nathan Smith, Tessa Davey, Timothy Lichtenstein, and Timothy Pillsbury. Without all of you in my life I do not know what I would do. Graduate school can be difficult but with you guys I enjoyed it every step of the way. I found a family here with all of you and became part of a group of misfit toys, nomad chemistry and a cohort. You taught me new things, explored new places with me, and served as constant support. I will never be able to express how thankful I am to have you all in my life and how fondly I will look back at all of the things we did during this time. I am truly looking forward to all of our future adventures.