% Place your vita below.

Cassie Marker was born in Lansing, Michigan and after graduating from Haslett High School in 2008, Cassie enrolled at Alma College where she studied Chemistry and Math. After graduation she worked as a research technician for the Dow Chemical Company. Cassie then joined Dr. Zi-Kui Liu's research group at the Pennsylvania State University in 2013. During her graduate studies, she has had the opportunity to work at ATI specialty materials and South Central University in China. She won the U.S. DOE Graduate Student Research Award. Cassie is a NSF National Research Trainee Fellow and obtained a graduate minor in Computational Sciences. 
\newline
\newline
\noindent Listed below are her publications during her Ph.D. study:
\begin{enumerate}
	\item C. Marker, A. Ross, and Z.-K. Liu, "Chapter 3 Design of Materials Processing Using Computational Thermodynamics" submitted to the book Computational Materials System Design by D. Shin and J. Saal to be published in 2017
	\item C. Marker, S. Shang, X. L. Liu, G. Lindwall, and Z.-K. Liu, "First-principles calculations and thermodynamic modeling of the Sn-Ta system," Calphad, vol. 57, pp. 46-54, Jun. 2017.
	\item Z. Liu, M. D. Biegalski, S. L. Hsu, S. L. Shang, C. Marker, J. Liu, L. Li, L. Fan, T. L. Meyer, A. T. Wong, J. A. Nichols, D. Chen, L. You, Z. Chen, K. Wang, K. Wang, T. Z. Ward, Z. Gai, H. N. Lee, A. S. Sefat, V. Lauter, Z. K. Liu, H. M. Christen, Epitaxial Growth of Intermetallic MnPt Films on Oxides and Large Exchange Bias. Adv. Mater., vol. 28, no. 1, pp. 118-123, 2016.
	\item C. Marker, S. L. Shang, J.-C. Zhao, Z. -K. Liu, "Effects of alloying elements on the elastic properties of bcc Ti-X alloys from first-principles calculations" submitted Journal of Alloys and Compounds.
	\item C. Marker, S. L. Shang, J.-C. Zhao, Z. -K. Liu, "Effects of alloying elements on the elastic properties of bcc Ti-X-Y alloys from first-principles calculations" submitted Journal of Computational Materials Science
	\item T. Frueh, C. Marker, E. R. Kupp, C. Compson, J. Atria, J. Gray, Z. -K. Liu, and G. L. Messing, "Effects of Powder Chemistry on the Sintering of MgO-doped Bayer Al$_{2}$O$_{3}$" submitted Journal of American Ceramic Society
\end{enumerate}

